### 2012-03-23 figs/results.tex

% latex table generated in R 2.14.0 by xtable 1.7-0 package
% Thu Mar 22 18:27:09 2012
\begin{table}[t]
\begin{center}
{\footnotesize
\begin{tabular}{lllrrrrrrrr}
  \hline
dataset & likelihood & type & uniform & marg & online & K=1 & K=2 & K=3 & dp & truth \\ 
  \hline
synthetic & rem & train & -0.928 & -0.916 & -0.689 & -0.577 & -0.106 &  &  & -0.107 \\ 
   &  & test & -0.919 & -0.919 & -0.636 & -0.589 & -0.130 &  &  & -0.122 \\ 
   & mult & train & -4.605 & -4.593 & -4.281 & -4.157 & -3.688 &  &  & -3.689 \\ 
   &  & test & -4.605 & -4.606 & -4.225 & -4.179 & -3.729 &  &  & -3.720 \\ 
  eckmann-small & rem & train & -8.800 & -7.722 & -7.309 & -6.552 & -6.350 & -6.430 &  &  \\ 
   &  & test & -8.761 & -7.727 & -6.659 & -6.151 & -6.106 & -6.455 &  &  \\ 
   & mult & train & -8.955 & -7.889 & -7.462 & -6.514 & -6.438 & -6.456 &  &  \\ 
   &  & test & -8.955 & -7.933 & -6.849 & -6.051 & -6.238 & -6.340 &  &  \\ 
   \hline
\end{tabular}
}
\caption{Comparing mean loglikelihood for each event across methods for each dataset.  Larger values are better.  See text for details. [TODO: More datasets as well as results for the DP version.  Could also remove training set results.]}
\label{tab:results}
\end{center}
\end{table}