### 2012-05-30 docs/submission.tex

\documentclass{article}
\usepackage{nips12submit_e,times}
\usepackage{tikz}
\tikzstyle{vertex}=[auto=left,circle,fill=black!25,minimum size=20pt,inner sep=0pt]
\usepackage{verbatim}
\usepackage{url}
\usepackage{graphicx, color}
\usepackage{amsmath}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{footnote}

\title{Stochastic Blockmodels for Relational Event Data}
% Latent Variable Modeling of Relational Event Dynamics
% Latent Variable Modeling of Relational Event Sequences
% Latent Variable Modeling for Relational Event Sequences

\author{
David S.~Hippocampus\thanks{ Use footnote for providing further information
about author (webpage, alternative address)---\emph{not} for acknowledging
funding agencies.} \\
Department of Computer Science\\
Cranberry-Lemon University\\
Pittsburgh, PA 15213 \\
\texttt{hippo@cs.cranberry-lemon.edu} \\
\And
Coauthor \\
Affiliation \\
Address \\
\texttt{email} \\
\AND
Coauthor \\
Affiliation \\
Address \\
}

\begin{document}
\maketitle
\savenotes
\input{abstract}
\input{intro}
\input{model}
\input{inference}
\input{simulation}
\input{experiments}
\input{discussion}
%\subsubsection*{Acknowledgements}
\bibliographystyle{unsrt}
\bibliography{refs}
\input{appendices}
\spewnotes
\end{document}