### 2012-05-31 docs/submission.tex

\documentclass{article}
\usepackage{nips12submit_e,times}
\usepackage{tikz}
\tikzstyle{vertex}=[auto=left,circle,fill=black!25,minimum size=20pt,inner sep=0pt]
\usepackage{verbatim}
\usepackage{url}
\usepackage{graphicx, color}
\usepackage{amsmath}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{footnote}

\title{Stochastic blockmodeling of relational event dynamics}
% Dynamic stochastic equivalence for relational events
% Stochastic Blockmodels for Relational Event Data}
% Latent Variable Modeling of Relational Event Dynamics
% Latent Variable Modeling of Relational Event Sequences
% Latent Variable Modeling for Relational Event Sequences

% \author{
% Christopher DuBois
% Department of Statistics\\
% University of California, Irvine\\
% \texttt{duboisc@ics.uci.edu} \\
% \And
% Coauthor \\
% Affiliation \\
% Address \\

% \AND
% Coauthor \\
% Affiliation \\
% Address \\
% }

\begin{document}
\maketitle
\savenotes
\input{abstract}
\input{intro}
\input{model}
\input{inference}
\input{simulation}
\input{experiments}
\input{discussion}
%\subsubsection*{Acknowledgements}
\bibliographystyle{unsrt}
\bibliography{refs}
\input{appendices}
\spewnotes
\end{document}