### 2012-10-30 docs/submission.tex
Add AISTATS style files.

\documentclass[twoside]{article}
\usepackage{aistats2e}
% \documentclass{article}
% \usepackage{nips12submit_e,times}
\usepackage{tikz}
\tikzstyle{vertex}=[auto=left,circle,fill=black!25,minimum size=20pt,inner sep=0pt]
\usepackage{verbatim}
\usepackage{url}
\usepackage{graphicx, color}
\usepackage{amsmath}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{footnote}
\usepackage{sidecap}

%\usepackage[accepted]{aistats2e}

\begin{document}

\twocolumn[

\aistatstitle{Stochastic blockmodeling of relational event dynamics}

\aistatsauthor{ Anonymous Author 1 \And Anonymous Author 2 \And Anonymous Author 3 }

\aistatsaddress{ Unknown Institution 1 \And Unknown Institution 2 \And Unknown Institution 3 } ]

%\maketitle
\input{abstract}
\input{intro}
\input{model}
\input{inference}
\input{simulation}
\input{experiments}
\input{discussion}
%\subsubsection*{Acknowledgements}
\bibliographystyle{unsrt}
\bibliography{refs}
%\newpage
%\input{appendices}
\end{document}